---
title: "County Typology"
author: "(Your name here)"
date: 2020-
output:
  github_document:
    toc: true
---

*Purpose*: In this challenge, you'll learn how to navigate the U.S. Census Bureau website, programmatically download data from the internet, and perform a county-level analysis. This will give you a base on which we'll build for Project 1.

<!-- include-rubric -->
# Grading Rubric
<!-- -------------------------------------------------- -->

Unlike exercises, **challenges will be graded**. The following rubrics define how you will be graded, both on an individual and team basis.

## Individual
<!-- ------------------------- -->

| Category | Needs Improvement | Satisfactory |
|----------|----------------|--------------|
| Effort | Some task __q__'s left unattempted | All task __q__'s attempted |
| Observed | Did not document observations, or observations incorrect | Documented correct observations based on analysis |
| Supported | Some observations not clearly supported by analysis | All observations clearly supported by analysis (table, graph, etc.) |
| Assessed | Observations include claims not supported by the data, or reflect a level of certainty not warranted by the data | Observations are appropriately qualified by the quality & relevance of the data and (in)conclusiveness of the support |
| Code Styled | Violations of the [style guide](https://style.tidyverse.org/) hinder readability | Code sufficiently close to the [style guide](https://style.tidyverse.org/) |

## Team
<!-- ------------------------- -->

| Category | Needs Improvement | Satisfactory |
|----------|----------------|--------------|
| Documented | No team contributions to Wiki | Team contributed observations to Wiki |
| Referenced | No team references in Wiki | At least one reference in Wiki to member report(s) |
| Relevant | References unrelated to observations, or difficult to find related analysis based on reference text | Reference text clearly points to relevant analysis with active hyperlinks |

## Due Date
<!-- ------------------------- -->

All the deliverables stated in the rubrics above are due **at midnight** before the day of the class discussion of the challenge. See the [Syllabus](https://docs.google.com/document/d/1jJTh2DH8nVJd2eyMMoyNGroReo0BKcJrz1eONi3rPSc/edit?usp=sharing) for more information.

```{r setup}
library(tidyverse)

```

*Background*: The USDA publishes [county typology codes](https://www.ers.usda.gov/data-products/county-typology-codes.aspx) to classify US counties to provide policy-relevant data. We will study their 2015 data in this challenge.

# The Big Picture
<!-- -------------------------------------------------- -->

We're about to go through *a lot* of weird steps, so let's first fix the big picture firmly in mind:

We want to study US counties in terms population and USDA typology. These data come from different sources, so we'll need to clean and join two datasets together.

That's the high-level view; now let's dig into the details.

# Get the Data
<!-- -------------------------------------------------- -->

1. County-level population estimates (Census Bureau)
2. County-level typology codes

## Navigating the Census Bureau
<!-- ------------------------- -->

**Steps**: Our objective is to find the 2018 American Community Survey[1] (ACS) Total Population estimates, disaggregated by counties. To check your results, this is Table `B01003`.

1. Go to [data.census.gov](data.census.gov).
2. Scroll down and click `View Tables`.
3. Apply filters to find the ACS **Total Population** estimates, disaggregated by counties. I used the filters:
  - `Topics > Populations and People > Counts, Estimates, and Projections > Population Total`
  - `Geography > County > All counties in United States`
5. Select the **Total Population** table and click the `Download` button to download the data; make sure to select the 2018 5-year estimates.
6. Unzip and move the data to your `challenges/data` folder.
  - Note that the data will have a crazy-long filename `ACSDT5Y2018.B01003_data_with_overlays_2020-07-26T094857.csv`. That's because metadata is stored in the filename, such as the year of the estimate (`Y2018`) and my access date (`2020-06-30`). **Your filename will vary based on when you download the data**, so make sure to copy the filename that corresponds to what you downloaded!

__q1__ Load Table `B01003` into the following tibble. Make sure the column names are `id, Geographic Area Name, Estimate!!Total, Margin of Error!!Total`.

*Hint*: You will need to use the `skip` keyword when loading these data!

```{r q1-task}
## TASK: Load the census bureau data with the following tibble name.
df_pop <- NA
```

*Note*: You can find information on 1-year, 3-year, and 5-year estimates [here](https://www.census.gov/programs-surveys/acs/guidance/estimates.html). The punchline is that 5-year estimates are more reliable but less current.

## Automated Download of USDA Typology Data
<!-- ------------------------- -->

__q2__ Head to the [USDA Typology](https://www.ers.usda.gov/data-products/county-typology-codes.aspx) site and copy the link for `ERS County Typology Codes, 2015 Edition (in CSV format)`. Paste that link as a string below, and assign it to `url_counties`.

```{r q2-task}
## TASK: Find the URL for the NYT covid-19 county-level data
url_counties <- NA_character_
```

Once you have the url, the following code will download a local copy of the data, then load the data into R.

```{r download}
## NOTE: No need to change this; just execute
## Set the filename of the data to download
filename_usda <- "./data/usda_codes.csv"

## Download the data locally
curl::curl_download(
        url_counties,
        destfile = filename_usda
      )

## Loads the downloaded csv
df_usda <-
  read_csv(
    filename_usda,
    col_types = cols(
      FIPStxt = col_character(),
      State = col_character(),
      County_name = col_character(),
      `Metro-nonmetro status, 2013 0=Nonmetro 1=Metro` = col_logical(),
      `Economic Types Type_2015_Update non-overlapping` = col_double(),
      Economic_Type_Label = col_character(),
      Farming_2015_Update = col_double(),
      `Mining_2015-Update` = col_double(),
      Manufacturing_2015_Update = col_double(),
      Government_2015_Update = col_double(),
      Recreation_2015_Update = col_double(),
      Nonspecialized_2015_Update = col_double(),
      Low_Education_2015_Update = col_double(),
      Low_Employment_Cnty_2008_2012_25_64 = col_double(),
      Pop_Loss_2010 = col_double(),
      Retirement_Dest_2015_Update = col_double(),
      Persistent_Poverty_2013 = col_double(),
      Persistent_Related_Child_Poverty_2013 = col_double()
    )
  ) %>%
  ## Fix the FIPS code
  mutate(
    fips = if_else(str_length(FIPStxt) == 4, str_c("0", FIPStxt), FIPStxt)
  ) %>%
  select(-FIPStxt) %>%
  select(fips, everything())
```

You can now re-run the chunk above (or the entire notebook) to pull the most recent version of the data. The USDA data are not changing over time, but you can re-use this for data sources that are updating.

*Note*: You should feel free to copy-paste the code above for your own future projects!

# Join the Data
<!-- -------------------------------------------------- -->

To get a sense of our task, let's take a glimpse at our two data sources.

```{r glimpse}
## NOTE: No need to change this; just execute
df_pop %>% glimpse
df_usda %>% glimpse
```

To join these datasets, we'll need to use [FIPS county codes](https://en.wikipedia.org/wiki/FIPS_county_code).[2] The last `5` digits of the `id` column in `df_pop` is the FIPS county code, while the NYT data `df_covid` already contains the `fips`.

__q3__ Process the `id` column of `df_pop` to create a `fips` column.

```{r q3-task}
## TASK: Create a `fips` column by extracting the county code
df_q3 <- NA
```

Use the following test to check your answer.

```{r q3-tests}
## NOTE: No need to change this
## Check known county
assertthat::assert_that(
              (df_q3 %>%
              filter(str_detect(`Geographic Area Name`, "Autauga County")) %>%
              pull(fips)) == "01001"
            )
print("Very good!")
```

Once you've extracted the county fips code, you can join the datasets.

__q4__ Join `df_usda` with `df_q3` by the `fips` column. Use the proper type of join to preserve all rows in `df_usda`.

```{r q4-task}
## TASK: Join df_usda and df_q3 by fips.
df_q4 <- NA
```

For convenience, I create some more convenient column names.

```{r rename}
## NOTE: No need to change; run this to produce a more convenient tibble
df_data <-
  df_q4 %>%
  select(
    state = State,
    county = County_name,
    fips,
    type = Economic_Type_Label,
    metro = `Metro-nonmetro status, 2013 0=Nonmetro 1=Metro`,
    population = `Estimate!!Total`,
    everything()
  )

df_data
```

# Analyze
<!-- -------------------------------------------------- -->

Now that we've done the hard work of loading and wrangling the data, we can finally start our analysis. First you'll go through a couple guided-EDA questions, then you'll be free to pose and answer your own questions.

__q5__ Visualize the distribution of county population for metro and non-metro areas. Answer the questions under *observations* below.

```{r q5-task}
## TODO: Visualize by freqpoly the population of metro and non-metro counties
```

**Observations**:

- Are there more metro or non-metro counties?
- Which---metro or non-metro---tend to be larger in population?
- How much overlap is there in population between the two categories?

__q6__ Inspect the next plot and document your observations.

```{r q6-task}
## NOTE: No need to edit; inspect and document your observations
df_data %>%

  ggplot(aes(fct_reorder(type, population), population)) +
  geom_boxplot(notch = TRUE) +

  coord_flip() +
  scale_y_log10() +
  labs(x = "County Type", y = "County Population")
```

**Observations**:

- What county types tend to be smaller (in population)?
- What county types tend to be larger (in population)?

__q7__ Drive your own ship: You've just put together a very rich dataset; you now get to explore! Pick your own direction and generate at least one punchline figure to document an interesting finding. I give a couple tips & ideas below:

### Ideas
<!-- ------------------------- -->

- Study some of the other variables, e.g. the education, unemployment, or poverty indicators.
- Look for outliers, e.g. particularly large or small counties of a given type.
- Do a websearch on outlier counties to learn more.
- Do a deep-dive on counties that are important to you (e.g. where you or your family live).
- Your own idea!

# Challenge Summary
<!-- -------------------------------------------------- -->

When you have completed this challenge, please write a summary about your findings and [submit it here](https://olin.qualtrics.com/jfe/form/SV_6sMhI2DEvNZBgXz). Remember that this summary is one of the graded components of the course!
